---
title: "Mini-Project 4: Text Analysis"
format:
  html: default
editor_options: 
  chunk_output_type: console
---

## Overview

You will find a data set containing string data.  This could be newspaper articles, tweets, songs, plays, movie reviews, or anything else you can imagine.  Then you will answer questions of interest and tell a story about your data using skills you have developed in strings, regular expressions, and text analysis.

Your story must contain the following elements:

- *at least* 3 different str_ functions
- *at least* 3 different regular expressions
- *at least* 2 different text analysis applications (count words, bing sentiment, afinn sentiment, nrc sentiment, wordclouds, trajectories over sections or time, tf-idf, bigrams, correlations, networks, LDA, etc.).  Note that many interesting insights can be gained by strategic and thoughtful use of regular expressions paired with simple counts and summary statistics.
- *at least* 3 illustrative, well-labeled plots or tables
- a description of what insights can be gained from your plots and tables.  *Be sure you weave a compelling and interesting story!*

Be sure to highlight the elements above so that they are easy for me to spot!


## Evaluation Rubric

Available [here](https://drive.google.com/file/d/1_yvxyKQrajvEOPZnZbriUDPdQ08DyCPG/view?usp=sharing)


## Timeline

Mini-Project 4 must be submitted on Moodle by **11:00 PM on Sun Apr 27**.  You should simply add a tab to your quarto webpage for Mini-Project 4, then you can just submit your URL (as long as your webpage also has a link to the GitHub repo containing your R code). 


## Topic Ideas

```{r}
#| include: false
#| message: false

library(lubridate)
library(tidyverse)
```


### Obama tweets

```{r}
#| message: false

#barack <- read_csv("Data/tweets_potus.csv") 
barack <- read_csv("https://joeroith.github.io/264_spring_2025/Data/tweets_potus.csv")
#michelle <- read_csv("Data/tweets_flotus.csv") 
michelle <- read_csv("https://joeroith.github.io/264_spring_2025/Data/tweets_flotus.csv")

tweets <- bind_rows(barack |> 
                      mutate(person = "Barack"),
                    michelle |> 
                      mutate(person = "Michelle")) |>
  mutate(timestamp = ymd_hms(timestamp))
```

President Barack Obama became the first US President with an official Twitter account, when @POTUS went live on May 18, 2015. (Yes, there was a time before Twitter/X.)  First Lady Michelle Obama got in on Twitter much earlier, though [her first tweet](https://youtu.be/3Pze1f1x7yU) was not from @FLOTUS. All of the tweets from @POTUS and @FLOTUS are now archived on Twitter as @POTUS44 and @FLOTUS44, and they are available as a csv download from the National Archive.  You can read more [here](https://obamawhitehouse.archives.gov/blog/2017/01/05/new-lenses-first-social-media-presidency).

Potential things to investigate:

- use of specific terms
- use of @, #, RT (retweet), or -mo (personal tweet from Michelle Obama)
- timestamp for date and time trends
- sentiment analysis
- anything else that seems interesting!


### Dear Abby advice column

Read in the “Dear Abby” data underlying The Pudding’s [30 Years of American Anxieties article](https://pudding.cool/2018/11/dearabby/).

```{r}
#| message: false

posts <- read_csv("https://raw.githubusercontent.com/the-pudding/data/master/dearabby/raw_da_qs.csv")
```

Take a couple minutes to scroll through the 30 Years of American Anxieties article to get ideas for themes that you might want to search for and illustrate using regular expressions.


### Other sources for string data

- [Other articles from The Pudding](https://github.com/the-pudding/data)
- NY Times headlines from the RTextTools package (see below)
- further analysis with the `bigspotify` data from class
- [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday)
- [kaggle](https://www.kaggle.com/datasets)
- [Data Is Plural](https://www.data-is-plural.com/)
- the options are endless -- be resourceful and creative!

```{r}
#| message: false

library(RTextTools)  # may have to install first
data(NYTimes)
as_tibble(NYTimes)
```

